Interactively Visualizing Multivariate Market Segmentation Using the R Package Lionfish

### 2024-10-02 manuscript.tex
a few minor adjustments. Expanded the Australian tourism applicaiton section to consider couples and families.

While tour animations are best obtained within \proglang{R} using the tourr package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}. Interactive graphics are available when using \proglang{Javascript}, as implemented in the \texttt{detourr} package \citep{RJ-2023-052}. This allows to replay a recorded tour path with interactive graphics, and can also be linked with additional displays, but lacks capabilities for manual tours.

To integrate user interactions with the capabilities of the \texttt{tourr package} an active communication with the interface is required. For example, we may wish to explore the local neighbourhood of a projection selected by the user with a local tour animation provided by tourr, or we may want to optimize a guided tour path using groups identified via brushing.
Our solution is using \proglang{Python} for high-performance interactivity, through the packages \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with integration to the tourr package \citep{tourr} via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between \proglang{Python} and \proglang{R}.

@@

The \texttt{lionfish} GUI can be launched using the function \texttt{interactive\textunderscore tour()}. At a minimum, the user needs to provide both the dataset and the instructions for constructing the desired plots. The dataset must be supplied as a \texttt{data.table}, while the plotting instructions should be passed as a list containing the named elements \texttt{type} and \texttt{obj}. The \texttt{type} element specifies the type of display to generate, such as \texttt{scatter} for a scatterplot or \texttt{2d\textunderscore tour} for a 2-dimensional tour. The \texttt{obj} element further defines the properties of the chosen display. For example, to create a 2-dimensional tour, the user must provide a \texttt{tour\textunderscore history} object, which can be generated using the \texttt{tourr} package. For a scatterplot, the user needs to provide a vector of strings specifying the names of the features to be displayed. The user can optionally specify the feature names, the arrangement of the plots, predefined subsets of the data (e.g., cluster solutions), custom names for these subsets, the number of available subsets, and the size of each plot.

@@

The \texttt{Save projections and subsets} button can be used to save the current state of the analysis. These states can be recovered by using pressing the \texttt{Load projections and subsets} button. Additionally, users can initiate new tours directly from the sidebar. The options for new tour paths are: a local tour around the currently shown projection and guided tours that search for projections based on the holes- or the linear discriminant analysis (LDA) index. The holes index is sensitive to projections with few points in the center of the projections and the LDA index aims to maximize the distance between the centers of the selectet subsets. \citep{ggobi} \textcolor{red}{XXX is this true? other options?} Additionally, all tour displays allow users to perform a manual tour; by right-clicking and dragging the arrowheads of projection axes, the projection is recalculated accordingly. This allows for manual exploration of the data.

\textcolor{red}{Mention additional interactive features in the displays - manual tour, anything else that might be relevant?}

@@

\textcolor{red}{Maybe here you can say something about adding new display types, what would be needed for implementing them}
\textcolor{blue}{I am unsure if this can easily be generalized. Some plots are easy to implement (scatterplot), but others like the mosaic plot come with a lot of complications, since that required a deep understanding on how matplotlib works. Implementing new interactive features would be especially challenging.}

@@

The second, $f^c$ is normalized relative to the size of each cluster and can be considered the distribution of features in each cluster. It is useful for comparing the composition of each cluster relative to the features. For example, for $p=4$, if $f^c_{1j} = (0.9, 0.2, 0.1, 0.1)$ it suggests that high values (1's) of feature 1 distinguishes cluster 1, and that values of the other  features were low. In the context of the later examples, this would mean cluster 1 contains tourists that especially engaged in activity 1, but not in activities 2, 3 and 4. This normalisation produces what is called the {\bf intra-cluster fraction}. \textcolor{red}{It corresponds to the cluster means. XXX why not larger values in the example, the maximum would be 1 so 0.4 still seems very moderate.}
\textcolor{blue}{I adjusted the example a bit.}

The last, $f^f$ is normalized relative to each feature, which can be considered to be the distribution of clusters on each feature. It is useful to examine how features are related to a cluster. Considering the last metric, for example, if $f^f_{i1} = (0.2, 0.7, 0, 0.1, 0)$ ($k=5$) would indicate that high values on feature 1 primarily are in cluster 2. In the context of the later examples, this would mean activity 1 is most commonly listed in cluster 2. This normalisation produces what is called the {\bf intra-feature fraction}. \textcolor{red}{Note that this must be interpreted with care in case of inbalanced cluster sizes.}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset, with manually adjusted projections. Projection axes with a norm < 0.1 have been blend out to reduce cluttering. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. Changing the projection axes of ``going to a spa'', ``museums'', ``sightseeing'' and ``using health facilities'' reveals the preferences and overlap of clusters 1 and 3.}
    \label{fig:winter_cl7_pre}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_post.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset, after new subsets have been selected. Projection axes with a norm < 0.1 have been blend out to reduce cluttering. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. We can see that now almost all museum goers are in the new manually selected subset 7 and what their preferences are compared to the other clusters.}
    \label{fig:winter_cl7_post}
\end{figure}

@@

\subsubsection{Segmentation of solo travellers}

In the heatmap displaying the intra-cluster fraction shown in Figure \ref{fig:aus_heatmapg}, we observe that subsets 1, 2, and 6 tend to prefer travelling alone. As a tourist agency, we might be interested in targeting solo travellers more effectively. However, subsets 2 and 6 also include individuals who enjoy spending time with their friends. To further explore the dataset, we can launch an \texttt{interactive\_tour()} with the same configuration as in the previous example. To achieve better separation between the subsets, we can skip to the last frames of a 2 dimensional tour optimized for the LDA index.

Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis ``Friends'' outward to separate the data based on that feature. Datapoints in the opposite direction of the ``Friends'' axes are the solo travellers we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot. Subsequently, we can pull all other projection axes in one direction to separate data points based on their overall activity level. Tourists that fall in the direction of the axes generally engaged in more activities compared to those in the opposite direction of the projection axes. This separation is evident, as observations in subset 6 (brown) are located opposite the axes, and the heatmap shows that they did not engage in many activities. Similarly we can also see that subset 2 (orange), which contains quite active tourists, is shifted towards the direction of the axes.

@@

By comparing subset 3 (green), which contains active tourists who spent time with their friends, with subset 7 (pink), the active tourists who traveled alone, we notice that subset 7 showed less interest in visiting the casino, theatre, and chartering a boat.

We also observe distinctive differences in the interests of the subsets that traveled alone. Subset 7 was interested in relaxing, shopping, sightseeing, wildlife, going to the beach, visiting pubs, and exploring farms. Given that subset 7 showed a notable interest in going to pubs, we can assume they want to meet new people. This insight can be utilized in a marketing campaign that bundles the activities subset 7 is interested in, creating a package tailored for solo travellers. This way, they can meet other solo travellers while engaging in activities they enjoy.

Although subset 8 (grey) was generally less active, almost everyone still engaged in sightseeing. For them, the focus was on sightseeing, relaxing, shopping, and going to the beach, with much less interest in other activities. It can be assumed that they value their time alone. These tourists could potentially be targeted more effectively by offering sightseeing options with minimal interaction, such as using a phone app to provide information about interesting locations, rather than relying on a tour guide.

Finally, since subset 9 (gold) is notably inactive, one might infer that they prefer spending much of their time in their accommodation. Consequently, these individuals might be most interested in accommodations that offer well-equipped, comfortable living spaces with amenities that cater to relaxation and leisure.

@@

Similarly, in the Australian dataset, manual exploration allowed us to refine the understanding of solo travellers by dividing them into three distinct subsegments: highly active, moderately active, and largely inactive tourists. This nuanced segmentation, derived from a blend of automated clustering and manual adjustments, offers a deeper understanding of the varied needs and behaviors within this group, enabling more effective marketing strategies.

@@

Future developments in the software might include expanding the range of interactive features and providing additional visualization methods. Because it is built on \proglang{R} and \proglang{Python}, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI. For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours~\citep{Laa2020}, would be recommended. Furthermore, it provides an additional demonstration of how \proglang{R} and \proglang{Python} can be integrated to harness the strengths of both programming languages.

>>>>

While tour animations are best obtained within \proglang{R} using the \texttt{tourr} package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}. Interactive graphics are available when using \proglang{Javascript}, as implemented in the \texttt{detourr} package \citep{RJ-2023-052}. This allows to replay a recorded tour path with interactive graphics, and can also be linked with additional displays, but lacks capabilities for manual tours.

To integrate user interactions with the capabilities of the \texttt{tourr} package an active communication with the interface is required. For example, we may wish to explore the local neighbourhood of a projection selected by the user with a local tour animation provided by \texttt{tourr}, or we may want to optimize a guided tour path using groups identified via brushing.
Our solution is using \proglang{Python} for high-performance interactivity, through the packages \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with integration to the \texttt{tourr} package \citep{tourr} via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between \proglang{Python} and \proglang{R}.

@@

The \texttt{lionfish} GUI can be launched using the function \texttt{interactive\textunderscore tour()}. At a minimum, the user needs to provide both the dataset and the instructions for constructing the desired plots. The dataset must be supplied as a numeric matrix, while the plotting instructions should be passed as a list containing the named elements \texttt{type} and \texttt{obj}. The \texttt{type} element specifies the type of display to generate, such as \texttt{scatter} for a scatterplot or \texttt{2d\textunderscore tour} for a 2-dimensional tour. The \texttt{obj} element further defines the properties of the chosen display. For example, to create a 2-dimensional tour, the user must provide a \texttt{tour\textunderscore history} object, which can be generated using the \texttt{tourr} package. For a scatterplot, the user needs to provide a vector of strings specifying the names of the features to be displayed. The user can optionally specify the feature names, the arrangement of the plots, predefined subsets of the data (e.g., cluster solutions), custom names for these subsets, the number of available subsets, and the size of each plot.

@@

The \texttt{Save projections and subsets} button can be used to save the current state of the analysis. These states can be recovered by using pressing the \texttt{Load projections and subsets} button. Additionally, users can initiate new tours directly from the sidebar. The options for new tour paths are: a local tour around the currently shown projection and guided tours that search for projections based on the holes- or the linear discriminant analysis (LDA) index. The holes index is sensitive to projections with few points in the center of the projections and the LDA index aims to maximize the distance between the centers of the selectet subsets \citep{ggobi}. Additionally, all tour displays allow users to perform a manual tour; by right-clicking and dragging the arrowheads of projection axes, the projection is recalculated accordingly. This allows for manual exploration of the data.

More information on all the features of the GUI can be found on \href{https://github.com/mmedl94/lionfish}{https://github.com/mmedl94/lionfish} and \href{https://mmedl94.github.io/lionfish/}{https://mmedl94.github.io/lionfish/}.

@@

@@

The second, $f^c$ is normalized relative to the size of each cluster and can be considered the distribution of features in each cluster. It is useful for comparing the composition of each cluster relative to the features. For example, for $p=4$, if $f^c_{1j} = (0.9, 0.2, 0.1, 0.1)$ it suggests that high values (1's) of feature 1 distinguishes cluster 1, and that values of the other  features were low. In the context of the later examples, this would mean cluster 1 contains tourists that especially engaged in activity 1, but not in activities 2, 3 and 4. This normalisation produces what is called the {\bf intra-cluster fraction}. In case of binary features it equates to the cluster means.

The last, $f^f$ is normalized relative to each feature, which can be considered to be the distribution of clusters on each feature. It is useful to examine how features are related to a cluster. Considering the last metric, for example, if $f^f_{i1} = (0.2, 0.7, 0, 0.1, 0)$ ($k=5$) would indicate that high values on feature 1 primarily are in cluster 2. In the context of the later examples, this would mean activity 1 is most commonly listed in cluster 2. This normalisation produces what is called the {\bf intra-feature fraction}. It also has to be noted that this metric is heavily influenced by the cluster sizes. Consider a feature that is uniformly distributed across all clusters, then the largest cluster would receive the highest intra-feature fraction.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset, with manually adjusted projections. Projection axes with a norm < 0.1 have been blend out to reduce cluttering. Top left: 2D tour with a manually adjusted projection. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with a manually adjusted projection. Bottom right: mosaic plot. Changing the projection axes of ``going to a spa'', ``museums'', ``sightseeing'' and ``using health facilities'' reveals the preferences and overlap of clusters 1 and 3.}
    \label{fig:winter_cl7_pre}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_post.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset, after new subsets have been selected. Projection axes with a norm < 0.1 have been blend out to reduce cluttering. Top left: 2D tour with a manually adjusted projection. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with a manually adjusted projection. Bottom right: mosaic plot. We can see that now almost all museum goers are in the new manually selected subset 7 and what their preferences are compared to the other clusters.}
    \label{fig:winter_cl7_post}
\end{figure}

@@

\subsubsection{Segmentation of travellers travelling without friends}

In the heatmap displaying the intra-cluster fraction shown in Figure \ref{fig:aus_heatmapg}, we observe that subsets 1, 2, and 6 tend to prefer travelling without friends. We can assume that this group consits of tourists travelling alone as couples or with their family. Additional features to further subsegment these groups might be interesting for future surveys. As a tourist agency, we might be interested in targeting these travellers more effectively. However, subsets 2 and 6 also include individuals who enjoy spending time with their friends. To further explore the dataset, we can launch an \texttt{interactive\_tour()} with the same configuration as in the previous example. To achieve better separation between the subsets, we can skip to the last frames of a 2 dimensional tour optimized for the LDA index.

Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis ``Friends'' outward to separate the data based on that feature. Datapoints in the opposite direction of the ``Friends'' axes are the tourists we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot. Subsequently, we can pull all other projection axes in one direction to separate data points based on their overall activity level. Tourists that fall in the direction of the axes generally engaged in more activities compared to those in the opposite direction of the projection axes. This separation is evident, as observations in subset 6 (brown) are located opposite the axes, and the heatmap shows that they did not engage in many activities. Similarly we can also see that subset 2 (orange), which contains quite active tourists, is shifted towards the direction of the axes.

@@

By comparing subset 3 (green), which contains active tourists who spent time with their friends, with subset 7 (pink), the active tourists who traveled without their friends, we notice that subset 7 showed less interest in visiting the casino, theatre, and chartering a boat.

We also observe distinctive differences in the interests of the subsets that traveled without friends. Subset 7 was interested in relaxing, shopping, sightseeing, wildlife, going to the beach, visiting pubs, and exploring farms. Given that Subset 7 showed a notable interest in going to pubs, we can assume that the solo travelers and couples in this subset are looking to meet new people. This insight could be leveraged in a marketing campaign by bundling activities that appeal to Subset 7, creating packages tailored to these tourists. Such packages would provide opportunities for them to connect with other travelers who share similar interests while enjoying their preferred activities. Additionally, some solo travelers and couples might prefer experiences or environments without children, allowing for the creation of adult-oriented packages. The remaining subgroup within Subset 7 consists of families. To better cater to this group, packages can be refined to consider that not all museums or festivals are equally suitable for children, enabling more family-friendly customizations.

Although subset 8 (grey) was generally less active, almost everyone still engaged in sightseeing. For them, the focus was on sightseeing, relaxing, shopping, and going to the beach, with much less interest in other activities. It can be assumed that solo travellers in this group value their time alone. This subgroup could potentially be targeted more effectively by offering sightseeing options with minimal interaction, such as using a phone app to provide information about interesting locations, rather than relying on a tour guide. For couples in this subset, the focus may be on enjoying quality time together in a relaxed, low-key environment. Marketing strategies could emphasize romantic sightseeing experiences, such as sunset tours or private beach spots. Additionally, offering couples' packages that include spa treatments, leisurely dining experiences, and personalized itineraries could resonate well with this group, allowing them to unwind and connect at their own pace. Families in subset 8 might prioritize activities that balance relaxation and light exploration, particularly in child-friendly settings. Sightseeing tours tailored to families, with engaging and educational content for children, could be a strong fit. Beach outings that offer safe, family-oriented areas or interactive experiences like sandcastle-building competitions could also be appealing. By curating packages that cater to both relaxation and gentle family activities, families can enjoy their vacation with minimal stress.

Finally, since subset 9 (gold) is notably inactive, one might infer that they prefer spending much of their time in their accommodation. Consequently, these tourists might be most interested in accommodations that offer well-equipped, comfortable living spaces with amenities that cater to relaxation and leisure.

@@

Similarly, in the Australian dataset, manual exploration allowed us to refine the understanding of tourists travelling without friends by dividing them into three distinct subsegments: highly active, moderately active, and largely inactive tourists. This nuanced segmentation, derived from a blend of automated clustering and manual adjustments, offers a deeper understanding of the varied needs and behaviors within this group, enabling more effective marketing strategies.

@@

Future developments in the software might include expanding the range of interactive features and providing additional visualization methods. Because it is built on \proglang{R} and \proglang{Python}, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI. For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours~\citep{Laa2020}, would be recommended. The integration of new plot types could also be streamlined to improve the versitility of the package.

Furthermore, \texttt{lionfish} provides an additional demonstration of how \proglang{R} and \proglang{Python} can be integrated to harness the strengths of both programming languages.
